---
title       : Car's Explorer
subtitle    : Shiny App - Developing Data Products Course Project
author      : Mariana Simons
job         : July 21, 2014
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Car's Explorer Essentials

### *Research question* 

The **Car's Explorer** Shiny application is created to help car's users to investigate the question: "Is an automatic or manual car transmission better for miles per gallon (mpg)?"

### *Data set*

To address this question the Shiny application uses the *mtcars* in  ``` library(datasets)```.

### *Application setup*

The **Car's Explorer** Shiny application has two parts:
 - Part I: **Explore car's mpg** allows the user to explore mpg vs. transmission or weight.
 - Part II: **Predict car's mpg** allows the user to predict mpg for user's choice of transmission and weight.

--- .class #id 

## Part I: Explore car's mpg (mpg ~ transmission)

### **User's input:** transmission (via selectBox)  
### **Shiny app's output:**  the box plot shown below

```{r boxplot, fig.height=6, fig.align='center', echo = FALSE}
library(datasets)
data(mtcars)
mtcars$am <- as.factor(mtcars$am)
plot(mtcars$am, mtcars$mpg, xlab = "transmission (am = 0 - automatic, am = 1 - manual)",
     ylab = "mpg", main = "Exploratory plot of mpg ~ transmission", col = "blue")
```

--- .class #id 

## Part I: Explore car's mpg (mpg ~ weight)

### **User's input:** weight (via selectBox)  
### **Shiny app's output:**  the scatter plot shown below

```{r scatterplot, fig.height=6, fig.align='center', echo = FALSE}
library(datasets)
data(mtcars)
mtcars$am <- as.factor(mtcars$am)
plot(mtcars$wt, mtcars$mpg, xlab = "weight (lb/1000)", ylab = "mpg", 
     main = "Exploratory plot of mpg ~ weight", col = "blue", bg = "blue", pch = 19)
```


--- .class #id 

## Part II: Predict car's mpg

### **User's input:** Example: transmission = manual and weight = 1.8
### **Shiny app's output:**  Linear regression model prediction + plot (blue arrow dashed line).
*Predicted mpg for your input of transmission and weight is: 29.94*

```{r fitplot, fig.height=6, fig.align='center', echo = FALSE}
fit <- lm(mpg ~ wt + am  + am*wt, data = mtcars)
transmission <- 1
weight <- 1.8
prediction <- coef(fit)[1] + coef(fit)[2]*weight + coef(fit)[3]*transmission + coef(fit)[4]*weight*transmission
      
plot(mtcars$wt,mtcars$mpg, pch=19, col=mtcars$am, xlab = "Weight (lb/1000)", ylab = "mpg")
abline(fit$coeff[1], fit$coeff[2],lwd=3, col = "black")
abline(fit$coeff[1] + fit$coeff[3], fit$coeff[2] + fit$coeff[4],lwd=3, col = "red")
lines(c(weight, weight), c(10, prediction), col = "blue", lwd = 3, lty = "dashed")
points(weight, prediction, col = "blue", pch = 25, bg = "blue", cex = 1.8)
title(" Linear regression mpg ~ transmission + weight + transmission*weight")
legend("topright", pch = 19, col = c("black","red"), legend = c("automatic transmission", "manual transmission"))
```

